---
title: "Programming in tidyverse"
format: pdf
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Task 1

### Part a

We cannot use `read_csv()` for the `data.txt` file because this data is separated by `;`, while `read_csv()` can only read in data with values that are separated by `,`. For the `data.txt` file, we can use `read_csv2()` to read in data that is separated by `;`.

```{r}
data <- read_csv2("data/data.txt", col_types = "d", skip = 0)
data
```

### Part b

```{r}
data2 <- read_delim("data/data2.txt", delim = "6", col_types = "fdc")
data2
```

## Task 2

### Part a

```{r}
trailblazer <- read_csv("data/trailblazer.csv")
trailblazer
```

### Part b

```{r}
trailblazer_longer <- trailblazer |>
  pivot_longer(cols = !Player,
               names_to = "Games_Location",
               values_to = "Points") |>
  separate_wider_delim(cols = Games_Location, 
                       delim = "_",
                       names = c("Game", "Location"),
                       cols_remove = TRUE)
trailblazer_longer %>% slice_head(n = 5)
```

###Part c

```{r}
trailblazer_wider <- trailblazer_longer |> 
  pivot_wider(names_from = Location,values_from = Points) |> 
  group_by(Player) |>
  mutate(mean_home = mean(Home, na.rm = TRUE),
         mean_away = mean(Away, na.rm = TRUE))
trailblazer_wider
```

```{r}
trailblazer_wider |> group_by(Player) |>
  group_by(Player) |>
  mutate(mean_home = mean(Home, na.rm = TRUE),
         mean_away = mean(Away, na.rm = TRUE)) |>
  mutate(home_away_diff = mean_home - mean_away) |>
  arrange(desc(home_away_diff))
```

On average, Jusuf Nurkic scored more points at home than away through the first 10 games of the season as shown with the highest `home_away_diff`.

## Task 3

### Part a

-   \<NULL\> This is showing that there is not a value that exists for this variable.

-   \<dbl \[52\]\> This is showing that this variable contain 52 different values.

-   \<list\> This shows that the column contains lists as its values.

### Part b

```{r}
colleague_penguins <- penguins |> select(species, island, bill_length_mm) |> 
  group_by(species, island) |> 
  summarise(n = n(), .groups = "drop") |> 
  pivot_wider(names_from = island, values_from = n) |> 
  mutate(across(.cols = -(1:1), .fns = ~replace_na(., replace = 0)))
colleague_penguins
```

## Task 4

```{r}
penguins_na_rm <- penguins |> select(species, bill_length_mm) |>
  mutate(bill_length_mm = replace(bill_length_mm, 
                                  is.na(bill_length_mm) & species == "Adelie",
                                  26),
         bill_length_mm = replace(bill_length_mm, 
                                  is.na(bill_length_mm) & species == "Gentoo",
                                  30)) |> 
  arrange(bill_length_mm)
penguins_na_rm %>% slice_head(n = 10)
```

